### 2022-12-05 dissertation/appendix/datasets.tex
✨ (review): Spelling and grammar, derived publications fix

This appendix provides insight into the preparation and usage of datasets that were used as part of the empirical process. The section is structured as follows:

\begin{itemize}
	\item \textbf{Section \ref{app:datasets:overview}} provides an overview of the data processing that was executed on each dataset.

	\item \textbf{Section \ref{app:datasets:example}} provides an example data processing script for the iris dataset. The same approach was followed for all datasets.
\end{itemize}

@@

\section{Example}\label{app:datasets:example}

This section provides an example data processing script that was used for the iris dataset. All the datasets used in this dissertation follow the same data processing principles.

\includepdf[pages=-]{attachments/iris.pdf}

>>>>

This appendix provides insight into the preparation and usage of datasets that were used as part of the empirical process. Section \ref{app:datasets:overview} provides an overview of the data processing that was executed on each dataset.

@@

### 2022-12-05 dissertation/chapters/01-introduction.tex
✨ (final-review): Chapter 9 Done
✨ (review): Spelling and grammar, derived publications fix

\Acfp{FFNN} are specific types of \acp{ANN}~\cite{ref:reed:1999}. A popular field of focus for studying \acp{ANN} is the process by which these models are trained.  The most common way of training \acp{FFNN} is \index{supervised learning}\textit{supervised learning}, which is a training technique that involves exposing the \acp{ANN} with input data and comparing the produced output data to that of predefined target data. Training of \acp{ANN} is seen as an optimisation problem. The \acs{ANN} maintains a set of parameters, referred to as ``weights'' and ``biases''. A search algorithm known as a \index{heuristic}\textit{heuristic}~\cite{ref:pearl:1984} is used to assign optimal values to the parameters of the \acs{ANN}, such that a specified objective function is minimised. The research presented in this disseration focuses on the development of a specific type of heuristic, called a \acf{HH} to train \acp{FFNN} in a supervised learning approach.

@@

This section provides the reader with a brief summary of the research field and outlines key concepts that contribute to the topics discussed in this disseration.

@@

\citeauthor{ref:grobler:2015}~\cite{ref:grobler:2015} mentioned that \acp{HH} have been shown to solve a number of problems including bin-packing, examination timetabling, production scheduling, the traveling salesman problem, vehicle routing problem and many more.

@@

Careful, systematic selection is thus required to find and select the best heuristic to train \acp{FFNN}. In the past, researchers selected the best heuristic by trail and error. A set of heuristics and carefully selected hyper-parameters would be implemented, followed by an empirical test to evaluate the performance of each heuristic for a given problem domain~\cite{ref:pillay:2015}. In this way, researchers were able to determine which heuristics and which hyper-parameters performed well for different problems. However, this approach is problematic, because it is time-consuming and laborious.

@@

      \item \textbf{Appendix~\ref{app:symbols}} lists and defines the mathematical symbols used in this work, categorised according to the relevant chapter in which they appear.

>>>>

\Acfp{FFNN} are specific types of \acp{ANN}~\cite{ref:reed:1999}. A popular field of focus for studying \acp{ANN} is the process by which these models are trained.  The most common way of training \acp{FFNN} is \index{supervised learning}\textit{supervised learning}, which is a training technique that involves exposing the \acp{ANN} with input data and comparing the produced output data to that of predefined target data. Training of \acp{ANN} is seen as an optimisation problem. The \acs{ANN} maintains a set of parameters, referred to as ``weights'' and ``biases''. A search algorithm known as a \index{heuristic}\textit{heuristic}~\cite{ref:pearl:1984} is used to assign optimal values to the parameters of the \acs{ANN}, such that a specified objective function is minimised. The research presented in this dissertation focuses on the development of a specific type of heuristic, called a \acf{HH} to train \acp{FFNN} in a supervised learning approach.

@@

This section provides the reader with a brief summary of the research field and outlines key concepts that contribute to the topics discussed in this dissertation.

@@

\citeauthor{ref:grobler:2015}~\cite{ref:grobler:2015} mentioned that \acp{HH} have been shown to solve a number of problems including bin-packing, examination timetabling, production scheduling, the travelling salesman problem, vehicle routing problem and many more.

@@

Careful, systematic selection is thus required to find and select the best heuristic to train \acp{FFNN}. In the past, researchers selected the best heuristic by trial and error. A set of heuristics and carefully selected hyper-parameters would be implemented, followed by an empirical test to evaluate the performance of each heuristic for a given problem domain~\cite{ref:pillay:2015}. In this way, researchers were able to determine which heuristics and which hyper-parameters performed well for different problems. However, this approach is problematic, because it is time-consuming and laborious.

@@

### 2022-12-05 dissertation/chapters/02-artificial-neural-networks.tex
✨ (review): Spelling and grammar, derived publications fix

The \acs{AN} implements a nonlinear mapping from $\mathbb{R}^{I}$ to
$\mathbb{R}^{T}$, usually in the ranges $[0,1]$ or $[-1,1]$, depending on the
activation function used~\cite{ref:engelbrecht:2007} and is given
as

@@

Numerical data must be normalised and scaled appropriately for the \acs{AN}. In general, input data is normalised to a range that is appropriate for the \index{activation function}activation function being used by the \acs{AN}. Normalisation and scaling yields input data that is comparable to the output produced by the \acs{AN}. For the purposes of this disseration, the \index{min-max scaler}min-max scaler~\cite{ref:al:2006} and the \index{standard score scaler}standard score scaler~\cite{ref:jain:2005} are considered.

@@

The connection strength that synapses in the \acs{BN} have is modeled in the \acs{AN} as weight vectors of numerical values associated with each dimension of the input. Weights can either dampen or strengthen the input by some negative or positive numerical value. Changes in the weight associated with a feature changes the influence that that particular feature has on the predicted output. Finding the correct value for each weight, such that the \acs{AN} yields optimal output, is an optimisation problem~\cite{ref:thierens:1993}. Research has shown that weight initialisation plays an important role in the efficient training of \acp{ANN}~\cite{ref:thimm:1995}.

@@

\acp{FFNN} were the first and simplest type of \acp{ANN} developed~\cite{ref:schmidhuber:2015} and implement input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement \index{fully connected topology}fully connected topologies. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes. Depending on the optimisation algorithm used, error correction information can be ``backpropagated'' through the network. An illustration of a \acs{FFNN} is given in Figure~\ref{fig:ffnn} below.

@@

\index{overfitting}Overfitting describes a scenario where the trained \acs{FFNN} performs well on training data, but does not generalise well to never before seen data from the test set~\cite{ref:tetko:1995, ref:geron:2017}. Geron~\cite{ref:geron:2017} describes \index{overfitting}overfitting as the case where \acs{FFNN} is too complex relative to the noisiness of the training data.

@@

Underfitting describes a scenario where the \acs{FFNN} is not able to effectively learn the underlying structure of the training data~\cite{ref:tetko:1995, ref:geron:2017}. Geron~\cite{ref:geron:2017} describe underfitting as the case where the \acs{FFNN} is too simple relative to the underlying structure of the training data.\\
\\
There are two types of supervised learning algorithms based on when weights are updated~\cite{ref:engelbrecht:2007}. These include stochastic training and batch
training.

>>>>

The \acs{AN} implements a non-linear mapping from $\mathbb{R}^{I}$ to
$\mathbb{R}^{T}$, usually in the ranges $[0,1]$ or $[-1,1]$, depending on the
activation function used~\cite{ref:engelbrecht:2007} and is given
as

@@

Numerical data must be normalised and scaled appropriately for the \acs{AN}. In general, input data is normalised to a range that is appropriate for the \index{activation function}activation function being used by the \acs{AN}. Normalisation and scaling yields input data that is comparable to the output produced by the \acs{AN}. For the purposes of this dissertation, the \index{min-max scaler}min-max scaler~\cite{ref:al:2006} and the \index{standard score scaler}standard score scaler~\cite{ref:jain:2005} are considered.

@@

The connection strength that synapses in the \acs{BN} have is modelled in the \acs{AN} as weight vectors of numerical values associated with each dimension of the input. Weights can either dampen or strengthen the input by some negative or positive numerical value. Changes in the weight associated with a feature changes the influence that that particular feature has on the predicted output. Finding the correct value for each weight, such that the \acs{AN} yields optimal output, is an optimisation problem~\cite{ref:thierens:1993}. Research has shown that weight initialisation plays an important role in the efficient training of \acp{ANN}~\cite{ref:thimm:1995}.

@@

\acp{FFNN} were the first and simplest type of \acp{ANN} developed~\cite{ref:schmidhuber:2015} and implement input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement \index{fully connected topology}fully connected topologies. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes. Depending on the optimisation algorithm used, error correction information can be propagated backwards through the network. An illustration of a \acs{FFNN} is given in Figure~\ref{fig:ffnn} below.

@@

\index{overfitting}Overfitting describes a scenario where the trained \acs{FFNN} performs well on training data, but does not generalise well to never before seen data from the test set~\cite{ref:tetko:1995, ref:geron:2017}. \citeauthor{ref:geron:2017}~\cite{ref:geron:2017} describes \index{overfitting}overfitting as the case where \acs{FFNN} is too complex relative to the noisiness of the training data.

@@

Underfitting describes a scenario where the \acs{FFNN} is not able to effectively learn the underlying structure of the training data~\cite{ref:tetko:1995, ref:geron:2017}. \citeauthor{ref:geron:2017}~\cite{ref:geron:2017} describe underfitting as the case where the \acs{FFNN} is too simple relative to the underlying structure of the training data.\\
\\
There are two types of supervised learning algorithms based on when weights are updated~\cite{ref:engelbrecht:2007}. These include stochastic training and batch
training.
